---
title: "Shiny App"
format:
  html:
    toc: false
---

```{r}
shiny_app_url <- Sys.getenv("SHINY_APP_URL", unset = "http://127.0.0.1:3838")
is_local_target <- grepl("127\\.0\\.0\\.1|localhost", shiny_app_url)
shiny_target_label <- if (is_local_target) "Local target" else "Deployment target"
shiny_target_note <- if (is_local_target) {
  "The site currently points to a locally running Shiny process on port 3838."
} else {
  "The site is currently configured to point to the deployed Shiny application."
}
```

## Open The App

This page is the website entry point for the Shiny application. Right now it is configured to open:

`r shiny_app_url`

::: {.grid-cards}
::: {.card-block}
### Launch Shiny App

[Open the Shiny app](`r shiny_app_url`){target="_blank"}

`r shiny_target_note`
:::

::: {.card-block}
### Current Target

- Label: `r shiny_target_label`
- URL: `r shiny_app_url`
- Mode: `r if (is_local_target) "Local development" else "External deployment"`
:::

::: {.card-block}
### Run Command

```r
shiny::runApp("app")
```

Or from the project root:

```powershell
Rscript run_app.R 3838
```
:::
:::

## Embedded Preview

When the target app is available, the frame below gives the website a direct in-page preview.

```{r, echo=FALSE}
htmltools::div(
  style = "background:#fff;border:1px solid #e3d8c8;border-radius:18px;overflow:hidden;box-shadow:0 12px 28px rgba(27,41,45,0.06);",
  htmltools::div(
    style = "padding:0.9rem 1rem;border-bottom:1px solid #eadfce;background:#fffaf3;font-weight:700;",
    sprintf("Shiny Preview (%s)", shiny_app_url)
  ),
  htmltools::tags$iframe(
    src = shiny_app_url,
    title = "Singapore Tourism Time-Series VA Prototype",
    style = "width:100%;height:900px;border:0;background:#fff;",
    loading = "lazy"
  )
)
```

## Current Module Status

| Module | Current status | Current focus |
|---|---|---|
| Time Series Visual Analysis | active prototype shell | shared monthly trend reading and metadata lookup |
| Time Series Clustering | active prototype | Wang branch interaction, processed-data contract, focus-market narrative |
| Forecasting | active prototype | Jin branch country-arrivals forecasting with hotel/stay context |

## Data Contract

The shared project backbone for country-level arrivals uses:

- `data/raw/visitor_arrivals_full_dataset.xlsx`
- `data/processed/clustering_country_wide.csv`
- `data/processed/clustering_country_long.csv`
- `data/processed/clustering_series_metadata.csv`

The forecasting layer keeps the same country-arrivals backbone, then supplements interpretation with supporting tourism indicators from:

- `data/raw/Name your insight (4).xlsx`

Those supporting indicators are limited to:

- `Hotel Room Occupancy Rate`
- `Average Length of Stay`
- `Number of Hotels`
- `Total Room Revenue`

## Expected User Flow

1. Open the Shiny app from this page.
2. Enter the `Time Series Explorer` tab to inspect one monthly series and confirm its source and unit.
3. Enter the `Cluster Prototype` tab to compare country-arrivals recovery trajectories.
4. Enter the `Forecasting Prototype` tab to forecast one country-arrivals series and interpret it against hotel and stay indicators.

## Deployment Switch

For local testing, no extra configuration is needed because the page defaults to:

`http://127.0.0.1:3838`

When you are ready to switch to `shinyapps.io`, re-render the site with:

```powershell
$env:SHINY_APP_URL='https://YOUR-APP-NAME.shinyapps.io/YOUR-APP/'
quarto render
```

That will update both the launch button and the embedded preview to the deployed app.

## Notes For Local Testing

1. Start the app first if the preview frame is blank.
2. Use the standalone launch link if your browser blocks the embedded frame.
3. Keep the browser window at a normal desktop width when checking layout quality.
