---
title: "Singapore Tourism Recovery Visual Analytics EDA"
format: html
author: "Xi Zixun"
---

## Background

Singapore’s inbound tourism landscape is characterised by a concentrated set of major source markets and a geographical distribution that evolves over time. Analysis of visitor arrivals from 2017 to 2025 indicates that Singapore’s tourism demand is anchored primarily in the Asia-Pacific region, while the relative importance of individual source countries shifts across different years. Such variation is visible not only in country rankings and market shares, but also in the wider spatial pattern of visitor origins. These dynamics make it insufficient to rely on static summary figures alone. A more effective approach is to integrate geospatial and comparative visual analysis in order to examine both the global distribution of visitor origins and the changing prominence of leading markets. Accordingly, this project proposes an interactive visual analytics prototype to support a clearer understanding of where Singapore’s visitors come from and how the structure of its source markets changes over time.

## Preparation

### Package Installation & Loading

-   `tidyverse`- A family of modern R packages specially designed to support data science, analysis and communication task including creating static statistical graphs.
-   `plotly` - R library for plotting interactive statistical graphs.
-   `patchwork` - combining multiple ggplot2 graphs into one figure.
-   `readx1` - Graphical scales map data to aesthetics, and provide methods for automatically determining breaks and labels for axes and legends.
-   `lubridate` - provides intuitive functions for parsing, manipulating, and performing arithmetic on dates, times, and time zones

-   `cellranger` - Translate Spreadsheet Cell Ranges to Rows and Columns
-   `knitr` - Provides a general-purpose tool for dynamic report generation in R using Literate Programming techniques
-   `scales` - Graphical scales map data to aesthetics, and provide methods for automatically determining breaks and labels for axes and legends
```{r}
pacman::p_load(tidyverse, lubridate, plotly, patchwork, readxl, cellranger, knitr, scales)
```

## Data Exploration & Cleaning

```{r}
path <- "../data/raw/tourism_update.xlsx"

tour <- read_excel(
  path,
  sheet = "My Series",
  skip = 29,
  col_names = FALSE,
  .name_repair = "minimal"
)

names(tour) <- c(
  "Date",
  "Visitor Arrivals",
  "Tourism Receipts: YTD: YoY",
  "Tourist Expenditure Per Capita",
  "Visitor Arrivals: China",
  "Average Length of Stay",
  "Visitor Arrivals: Malaysia",
  "Visitor Arrivals: India",
  "Visitor Arrivals: Indonesia",
  "Visitor Arrivals: Australia",
  "Number of Hotels",
  "Visitor Days",
  "Hotel Room Occupancy Rate",
  "Visitor Arrivals: West Asia",
  "Visitor Arrivals: Taiwan",
  "Total Room Revenue",
  "Number of Hotels.1",
  "Average Length of Stay.1",
  "Visitor Arrivals: Hong Kong SAR (China)",
  "Hotel Revenue: per Available Room: Luxury",
  "Hotel Revenue: per Available Room: Mid-Tier",
  "Visitor Arrivals: ASEAN",
  "Visitor Arrivals: Italy",
  "Visitor Arrivals: Russian Federal (CIS)",
  "Visitor Arrivals: France",
  "Visitor Arrivals: Philippines",
  "Visitor Arrivals: Spain",
  "No of Hotel Room Stock",
  "Visitor Arrivals: Thailand",
  "Visitor Arrivals: Ireland",
  "Visitor Arrivals: United Arab Emirates",
  "Visitor Arrivals: United Kingdom",
  "Visitor Arrivals: Africa",
  "Visitor Arrivals: Bangladesh",
  "Visitor Arrivals: Iran",
  "Visitor Arrivals: New Zealand",
  "Visitor Arrivals: Israel",
  "Visitor Arrivals: North Asia",
  "Visitor Arrivals: 8-10 Days",
  "Visitor Arrivals: 11-14 Days",
  "Visitor Arrivals: 15 Days & Over",
  "Visitor Arrivals: Americas",
  "Visitor Arrivals: Germany",
  "Visitor Arrivals: Scandinavia: Sweden",
  "Visitor Arrivals: Switzerland",
  "Visitor Arrivals: USA",
  "Visitor Arrivals: Canada",
  "Visitor Arrivals: Mauritius",
  "Visitor Arrivals: Kuwait",
  "Visitor Arrivals: Egypt",
  "Visitor Arrivals: Brunei",
  "Visitor Arrivals: Finland",
  "Visitor Arrivals: Japan",
  "Visitor Arrivals: South Korea",
  "Visitor Arrivals: Myanmar",
  "Visitor Arrivals: Netherlands",
  "Visitor Arrivals: Scandinavia: Norway",
  "Visitor Arrivals: Saudi Arabia",
  "Visitor Arrivals: Sri Lanka",
  "Visitor Arrivals: Vietnam",
  "Visitor Arrivals: Pakistan",
  "Visitor Arrivals: Republic of South Africa"
)

tour <- tour %>%
  mutate(Date = as.Date(Date))

tour_clean <- tour %>%
  filter(year(Date) >= 2017, year(Date) != 2026)

shared_arrivals <- read.csv(
  "../data/processed/arrivals_country_long.csv",
  stringsAsFactors = FALSE,
  check.names = FALSE
) %>%
  mutate(
    date = as.Date(date),
    Country = country,
    Arrivals = arrivals,
    Year = year
  ) %>%
  filter(date >= as.Date("2017-01-01"), date <= as.Date("2025-12-01"))
```

```{r}
names(tour)
head(tour$Date)
class(tour$Date)
unique(year(tour_clean$Date))
```

::: callout-tip
Some preparation of data to be noticed:

-   The year 2015, 2016 and 2026 were removed because they do not consist of full year data from **Jan to Dec**
:::

```{r}
country_cols <- c(
  "Visitor Arrivals: China",
  "Visitor Arrivals: Malaysia",
  "Visitor Arrivals: India",
  "Visitor Arrivals: Indonesia",
  "Visitor Arrivals: Australia",
  "Visitor Arrivals: Taiwan",
  "Visitor Arrivals: Hong Kong SAR (China)",
  "Visitor Arrivals: Italy",
  "Visitor Arrivals: Russian Federal (CIS)",
  "Visitor Arrivals: France",
  "Visitor Arrivals: Philippines",
  "Visitor Arrivals: Spain",
  "Visitor Arrivals: Thailand",
  "Visitor Arrivals: Ireland",
  "Visitor Arrivals: United Arab Emirates",
  "Visitor Arrivals: United Kingdom",
  "Visitor Arrivals: Bangladesh",
  "Visitor Arrivals: Iran",
  "Visitor Arrivals: New Zealand",
  "Visitor Arrivals: Israel",
  "Visitor Arrivals: Germany",
  "Visitor Arrivals: Scandinavia: Sweden",
  "Visitor Arrivals: Switzerland",
  "Visitor Arrivals: USA",
  "Visitor Arrivals: Canada",
  "Visitor Arrivals: Mauritius",
  "Visitor Arrivals: Kuwait",
  "Visitor Arrivals: Egypt",
  "Visitor Arrivals: Brunei",
  "Visitor Arrivals: Finland",
  "Visitor Arrivals: Japan",
  "Visitor Arrivals: South Korea",
  "Visitor Arrivals: Myanmar",
  "Visitor Arrivals: Netherlands",
  "Visitor Arrivals: Scandinavia: Norway",
  "Visitor Arrivals: Saudi Arabia",
  "Visitor Arrivals: Sri Lanka",
  "Visitor Arrivals: Vietnam",
  "Visitor Arrivals: Pakistan",
  "Visitor Arrivals: Republic of South Africa"
)

tour_long <- shared_arrivals %>%
  transmute(
    Date = date,
    Country,
    Arrivals
  )
```

::: callout-note
The dataset was reshaped into long format so that each row represents one country’s visitor arrivals for one specific month. The resulting structure contains three main variables: `Date`, `Country`, and `Arrivals`. This format is more suitable for time-series visualization and comparative analysis in R, as it allows countries to be treated as categories within a single variable rather than as separate columns. As a result, it supports cleaner code, easier grouping and filtering, and more effective use of ggplot2 for visualizing changes in visitor arrivals over time.
:::

Before we continue with analysis results, some Date periods should be noticed: **AY2017 Start** - **AY2019 End** is consider as pre_covid period, **AY2020 Start** - **AY2021 End** is Consider as covid_shock, **AY2022 Star**t - **AY2025 End** is considered as recovery period

## EDA 1 - Singapore Visitors

```{r}
country_cols <- c(
  "Visitor Arrivals: China",
  "Visitor Arrivals: Malaysia",
  "Visitor Arrivals: India",
  "Visitor Arrivals: Indonesia",
  "Visitor Arrivals: Australia",
  "Visitor Arrivals: Taiwan",
  "Visitor Arrivals: Hong Kong SAR (China)",
  "Visitor Arrivals: Italy",
  "Visitor Arrivals: Russian Federal (CIS)",
  "Visitor Arrivals: France",
  "Visitor Arrivals: Philippines",
  "Visitor Arrivals: Spain",
  "Visitor Arrivals: Thailand",
  "Visitor Arrivals: Ireland",
  "Visitor Arrivals: United Arab Emirates",
  "Visitor Arrivals: United Kingdom",
  "Visitor Arrivals: Bangladesh",
  "Visitor Arrivals: Iran",
  "Visitor Arrivals: New Zealand",
  "Visitor Arrivals: Israel",
  "Visitor Arrivals: Germany",
  "Visitor Arrivals: Scandinavia: Sweden",
  "Visitor Arrivals: Switzerland",
  "Visitor Arrivals: USA",
  "Visitor Arrivals: Canada",
  "Visitor Arrivals: Mauritius",
  "Visitor Arrivals: Kuwait",
  "Visitor Arrivals: Egypt",
  "Visitor Arrivals: Brunei",
  "Visitor Arrivals: Finland",
  "Visitor Arrivals: Japan",
  "Visitor Arrivals: South Korea",
  "Visitor Arrivals: Myanmar",
  "Visitor Arrivals: Netherlands",
  "Visitor Arrivals: Scandinavia: Norway",
  "Visitor Arrivals: Saudi Arabia",
  "Visitor Arrivals: Sri Lanka",
  "Visitor Arrivals: Vietnam",
  "Visitor Arrivals: Pakistan",
  "Visitor Arrivals: Republic of South Africa"
)

tour_long <- shared_arrivals %>%
  transmute(
    Date = date,
    Country,
    Arrivals,
    Year
  )
```

```{r}
geo_year <- tour_long %>%
  group_by(Year, Country) %>%
  summarise(
    Arrivals = sum(Arrivals, na.rm = TRUE),
    .groups = "drop"
  )

geo_year <- geo_year %>%
  mutate(
    Country_map = recode(
      Country,
      "USA" = "United States",
      "Russian Federal (CIS)" = "Russia",
      "Hong Kong SAR (China)" = "Hong Kong",
      "Scandinavia: Sweden" = "Sweden",
      "Scandinavia: Norway" = "Norway",
      "Republic of South Africa" = "South Africa"
    )
  )

```

```{r}
plot_ly(
  data = geo_year,
  type = "choropleth",
  locations = ~Country_map,
  locationmode = "country names",
  z = ~Arrivals,
  frame = ~Year,
  text = ~paste(
    "Country:", Country,
    "<br>Year:", Year,
    "<br>Arrivals:", format(Arrivals, big.mark = ",")
  ),
  hoverinfo = "text",
  colorscale = "Reds"
) %>%
  layout(
    title = "Visitor Arrivals to Singapore by Source Country (2017-2025)",
    geo = list(showframe = FALSE, showcoastlines = TRUE)
  )

```

The geospatial results indicate that Singapore’s visitor arrivals are heavily concentrated in Asia, with nearby regional markets contributing the largest share of tourism inflows. This highlights the continued importance of geographical proximity, regional connectivity, and established travel links in shaping visitor demand. The map also shows that arrivals are not evenly distributed across the world, but are instead clustered around a limited number of key source markets.

Over time, the spatial pattern changes noticeably, especially around the pandemic period. Visitor distribution weakens during the years of disruption and gradually strengthens again in the later years, although the pace of recovery appears uneven across countries. This suggests that while Singapore’s tourism sector has recovered, the composition of its visitor markets may not have returned uniformly to its earlier pattern. Overall, the findings show that Singapore’s tourism demand remains regionally concentrated, but also sensitive to major external disruptions.

## EDA 2 - Top Visitor Countries Comparison Pre and Post Covid

```{r}
p1_data <- tour_long %>%
  filter(year(Date) >= 2017 & year(Date) <= 2019)

p2_data <- tour_long %>%
  filter(year(Date) >= 2022 & year(Date) <= 2025)
```

::: panel-tabset
## Table - Pre_Covid

```{r}
p1_total_market <- tour_clean %>%
  filter(year(Date) >= 2017 & year(Date) <= 2019) %>%
  summarise(
    Total_Market_Avg = round(mean(`Visitor Arrivals`, na.rm = TRUE), 0)
  ) %>%
  pull(Total_Market_Avg)

# Country ranking in P1
p1_rank <- p1_data %>%
  group_by(Country) %>%
  summarise(
    Avg_Monthly_Arrivals = round(mean(Arrivals, na.rm = TRUE), 0),
    .groups = "drop"
  ) %>%
  mutate(
    Overall_Share_Percent = (Avg_Monthly_Arrivals / p1_total_market) * 100
  ) %>%
  arrange(desc(Avg_Monthly_Arrivals)) %>%
  mutate(Rank = row_number()) %>%
  slice_head(n = 5)

p1_rank
```

## Chart - Pre_Covid

```{r}
plot_ly(
  data = p1_rank,
  x = ~Avg_Monthly_Arrivals,
  y = ~reorder(Country, Avg_Monthly_Arrivals),
  type = "bar",
  orientation = "h",
  text = ~paste0(round(Overall_Share_Percent, 1), "%"),
  textposition = "outside",
  hovertemplate = "Average Monthly Arrivals: %{x:,}<extra></extra>"
) %>%
  layout(
    title = "P1 (2017-2019): Top 5 Visitor Countries",
    xaxis = list(title = "Average Monthly Arrivals"),
    yaxis = list(title = "Country")
  )
```
:::

During the pre-COVID period (2017–2019), Singapore’s inbound tourism market was led by China, which recorded an average monthly arrival figure of 285,357 and accounted for 18.66% of Singapore’s total inbound visitor arrivals. This indicates that China was the most important single source market before the pandemic. India ranked second with 114,787 average monthly arrivals and an overall market share of 7.51%, followed by Malaysia at 101,198 (6.62%), Australia at 92,571 (6.05%), and Japan at 69,635 (4.55%).

These results suggest that Singapore’s pre-pandemic visitor market was concentrated in a relatively small number of major source countries, with China standing out very clearly as the dominant market. The gap between China and the other four countries is substantial, showing that the market was not evenly distributed even within the top five visitor sources. Instead, a significant share of total visitor demand was anchored in one especially dominant country, followed by a second tier of comparatively smaller but still important markets. This points to a tourism structure that was concentrated, but with a particularly strong dependence on China.

::: panel-tabset
## Table - Post_Covid

```{r}
p2_total_market <- tour_clean %>%
  filter(year(Date) >= 2022 & year(Date) <= 2025) %>%
  summarise(
    Total_Market_Avg = round(mean(`Visitor Arrivals`, na.rm = TRUE), 0)
  ) %>%
  pull(Total_Market_Avg)

# Country ranking in P2
p2_rank <- p2_data %>%
  group_by(Country) %>%
  summarise(
    Avg_Monthly_Arrivals = round(mean(Arrivals, na.rm = TRUE), 0),
    .groups = "drop"
  ) %>%
  mutate(
    Overall_Share_Percent = (Avg_Monthly_Arrivals / p2_total_market) * 100
  ) %>%
  arrange(desc(Avg_Monthly_Arrivals)) %>%
  mutate(Rank = row_number()) %>%
  slice_head(n = 5)

p2_rank
```

## Chart - Post_Covid

```{r}
plot_ly(
  data = p2_rank,
  x = ~Avg_Monthly_Arrivals,
  y = ~reorder(Country, Avg_Monthly_Arrivals),
  type = "bar",
  orientation = "h",
  text = ~paste0(round(Overall_Share_Percent, 1), "%"),
  textposition = "outside",
  hovertemplate = "Average Monthly Arrivals: %{x:,}<extra></extra>"
) %>%
  layout(
    title = "P2 (2022-2025): Top 5 Visitor Countries",
    xaxis = list(title = "Average Monthly Arrivals"),
    yaxis = list(title = "Country")
  )
```
:::

During the post-COVID recovery period (2022–2025), Singapore’s visitor market was led by China, which recorded an average monthly arrival figure of 159,896 and accounted for 14.39% of Singapore’s total inbound visitor arrivals. This indicates that China remained the most important single source market in the recovery phase, even though its scale was still below the pre-pandemic level. India ranked second with 86,675 average monthly arrivals and an overall market share of 7.80%, followed closely by Malaysia at 86,258 (7.76%), Australia at 85,258 (7.67%), and the Philippines at 53,741 (4.84%).

These results suggest that Singapore’s post-pandemic visitor structure remained concentrated in a limited number of major source markets, with China continuing to occupy a clearly dominant position. Although the ranking pattern changed compared with the pre-COVID period, the results still point to a market structure in which a small group of countries accounted for a substantial share of total visitor arrivals. At the same time, the gap between China and the remaining four countries remained large, showing that the recovery period did not eliminate the strong concentration of visitor demand at the top of the market.

A comparison with the pre-COVID baseline provides a clearer view of recovery status. China’s average monthly arrivals declined from 285,357 in P1 to 159,896 in P2, while India fell from 114,787 to 86,675, Malaysia from 101,198 to 86,258, and Australia from 92,571 to 85,258. The appearance of the Philippines in the top five, replacing Japan, suggests that the composition of Singapore’s key source markets shifted during the recovery period. This indicates that post-COVID recovery was not simply a return to the previous market structure, but involved some reordering in the relative importance of individual countries.


## Conclusion

Overall, the analysis shows that Singapore’s visitor market remained strongly concentrated in a relatively small number of major source countries across both the pre-COVID and post-COVID periods. In the pre-pandemic years, China clearly dominated Singapore’s inbound tourism market, followed by a smaller second tier of key Asia-Pacific markets. In the post-COVID period, China remained the leading source market, but at a lower scale than before, while the composition of the top five shifted slightly with the entry of the Philippines and the exit of Japan. This suggests that Singapore’s tourism recovery was only partial in terms of visitor volume, and that the structure of its key source markets also experienced some adjustment.

At the same time, the results consistently highlight Singapore’s strong dependence on Asia-Pacific markets. The leading visitor source countries in both periods were overwhelmingly drawn from Asia and the broader regional travel network, indicating that proximity, connectivity, and established travel demand remain central to Singapore’s tourism performance. Taken together, these findings suggest that while Singapore’s tourism market has shown resilience in maintaining a concentrated regional core, the recovery process has not fully restored the scale or exact composition of the pre-pandemic visitor structure.
