---
title: "Prototype: Time Series Clustering"
subtitle: "Country-level tourism recovery trajectories in Singapore"
format:
  html:
    fig-width: 9
    fig-height: 5.2
execute:
  echo: true
  warning: false
  message: false
---

## 1. Introduction

This prototype follows a strict code -> result -> explanation rhythm. Each section shows one piece of code, one visible result, and one interpretation.

The clustering unit in this module is a **country time series**, not a month. In other words, each country is treated as one monthly arrival trajectory, and the goal is to identify which countries share similar recovery patterns in Singapore's inbound tourism.

The page first looks for the processed files expected by the final project:

- `data/processed/clustering_country_wide.csv`
- `data/processed/clustering_country_long.csv`

If those files are not present during local rendering, the page falls back to the raw workbook so that the prototype remains executable while the project is being assembled.

The method used here is a practical hierarchical clustering workflow built on comparable monthly trajectories:

1. index each series to a common baseline;
2. z-score the indexed values across time for comparability;
3. cluster the country series with a distance matrix and Ward linkage;
4. use silhouette diagnostics to choose `k`;
5. interpret the final clusters with a membership table, representative profile plot, and a China placement check.

## 2. Load Packages

The first step is to load the packages used throughout the prototype.

```{r}
packages <- c("readxl", "dplyr", "ggplot2", "cluster", "knitr")
invisible(lapply(packages, library, character.only = TRUE))

package_table <- data.frame(
  Package = packages,
  Version = vapply(packages, function(pkg) as.character(packageVersion(pkg)), character(1))
)

knitr::kable(package_table, align = c("l", "l"))
```

The package table confirms that the prototype uses a small, base-project-friendly set of tools for reading data, shaping trajectories, clustering country series, and reporting results.

## 3. Import The Series Files

The next step is to import the processed clustering inputs if they already exist. If not, the page reconstructs a compact country-level series table from the raw workbook for local checks.

```{r}
find_project_root <- function(start = getwd()) {
  current <- normalizePath(start, winslash = "/", mustWork = TRUE)

  repeat {
    if (file.exists(file.path(current, "_quarto.yml"))) {
      return(current)
    }
    parent <- dirname(current)
    if (identical(parent, current)) {
      stop("Project root not found. Expected to locate _quarto.yml.")
    }
    current <- parent
  }
}

first_existing <- function(paths) {
  found <- paths[file.exists(paths)][1]
  if (length(found) == 0 || is.na(found) || !nzchar(found)) {
    NA_character_
  } else {
    found
  }
}

clean_series_name <- function(x) {
  x <- gsub("^Visitor Arrivals: ", "", x)
  x <- gsub("^\\s+|\\s+$", "", x)
  x
}

coerce_series_numeric <- function(df, cols) {
  for (col in cols) {
    df[[col]] <- as.numeric(gsub(",", "", df[[col]]))
  }
  df
}

project_root <- find_project_root()
analysis_start <- as.Date("2017-01-01")
analysis_end <- as.Date("2025-12-01")

wide_processed_path <- first_existing(c(
  file.path(project_root, "data", "processed", "clustering_country_wide.csv"),
  file.path("data", "processed", "clustering_country_wide.csv")
))
long_processed_path <- first_existing(c(
  file.path(project_root, "data", "processed", "clustering_country_long.csv"),
  file.path("data", "processed", "clustering_country_long.csv")
))
raw_workbook_path <- first_existing(c(
  file.path(project_root, "data", "raw", "visitor_arrivals_full_dataset.xlsx"),
  file.path("data", "raw", "visitor_arrivals_full_dataset.xlsx")
))

if (!is.na(wide_processed_path)) {
  country_wide <- read.csv(wide_processed_path, check.names = FALSE, stringsAsFactors = FALSE)
  country_wide$date <- as.Date(country_wide$date)
  country_wide <- country_wide |>
    dplyr::filter(date >= analysis_start, date <= analysis_end)
} else {
  raw_wide <- readxl::read_excel(raw_workbook_path, sheet = "My Series")
  raw_wide$date <- as.Date(raw_wide[[1]])
  raw_wide <- raw_wide |>
    dplyr::filter(date >= analysis_start, date <= analysis_end)

  fallback_series <- c(
    "Visitor Arrivals: China",
    "Visitor Arrivals: India",
    "Visitor Arrivals: Malaysia",
    "Visitor Arrivals: Australia",
    "Visitor Arrivals: Japan",
    "Visitor Arrivals: Philippines",
    "Visitor Arrivals: United Kingdom",
    "Visitor Arrivals: Germany",
    "Visitor Arrivals: Thailand",
    "Visitor Arrivals: Hong Kong SAR (China)"
  )
  available_series <- intersect(fallback_series, names(raw_wide))
  country_wide <- raw_wide |>
    dplyr::select(date, dplyr::all_of(available_series))
  names(country_wide) <- c("date", clean_series_name(names(country_wide)[-1]))
}

series_cols <- setdiff(names(country_wide), c("date", "year", "month", "quarter", "period"))
series_cols <- series_cols[vapply(country_wide[series_cols], is.numeric, logical(1))]
country_wide <- coerce_series_numeric(country_wide, series_cols)

if (!is.na(long_processed_path)) {
  country_long <- read.csv(long_processed_path, stringsAsFactors = FALSE)
  country_long$date <- as.Date(country_long$date)
  country_long <- country_long |>
    dplyr::filter(date >= analysis_start, date <= analysis_end)
} else {
  country_long <- data.frame(
    date = rep(country_wide$date, times = length(series_cols)),
    country = rep(series_cols, each = nrow(country_wide)),
    arrivals = as.vector(as.matrix(country_wide[series_cols]))
  )
  country_long$year <- as.integer(format(country_long$date, "%Y"))
  country_long$month <- as.integer(format(country_long$date, "%m"))
  country_long$indexed_arrivals <- ave(country_long$arrivals, country_long$country, FUN = function(x) x / x[1] * 100)
  country_long$zscore_arrivals <- ave(country_long$arrivals, country_long$country, FUN = function(x) as.numeric(scale(x)))
}

china_series <- series_cols[grepl("china", series_cols, ignore.case = TRUE)][1]
if (is.na(china_series) || !nzchar(china_series)) {
  stop("Unable to identify the China series in the clustering input.")
}

import_summary <- data.frame(
  Item = c(
    "Wide input",
    "Long companion",
    "Analysis window",
    "Number of countries",
    "Months",
    "China series"
  ),
  Value = c(
    ifelse(is.na(wide_processed_path), "Derived from raw workbook", wide_processed_path),
    ifelse(is.na(long_processed_path), "Derived from wide input", long_processed_path),
    paste(format(analysis_start, "%Y-%m"), "to", format(analysis_end, "%Y-%m")),
    length(series_cols),
    nrow(country_wide),
    china_series
  )
)

knitr::kable(import_summary, align = c("l", "l"))
```

The import summary confirms that the prototype works on a clean monthly window and that the clustering input is a compact country-level table rather than a month-level feature matrix.

## 4. Confirm The Clustering Unit

Before clustering, it is useful to inspect the country series that will be included in the model.

```{r}
country_summary <- country_long |>
  dplyr::group_by(country) |>
  dplyr::summarise(
    first_month = format(min(date), "%Y-%m"),
    last_month = format(max(date), "%Y-%m"),
    missing_arrivals = sum(is.na(arrivals)),
    mean_arrivals = round(mean(arrivals, na.rm = TRUE), 1),
    .groups = "drop"
  ) |>
  dplyr::arrange(dplyr::desc(mean_arrivals))

knitr::kable(country_summary, align = c("l", "l", "l", "r", "r"))
```

This table shows that each row is a country series with a common monthly span. That is exactly the right unit for time series clustering because the model is now grouping recovery trajectories instead of individual months.

## 5. Build Comparable Trajectories

The next step is to put every country onto a common scale. Indexing starts each series at 100, and z-scoring then makes the trajectories comparable for distance-based clustering.

```{r}
country_indexed <- country_wide
for (col in series_cols) {
  first_value <- country_wide[[col]][which(!is.na(country_wide[[col]]))[1]]
  country_indexed[[col]] <- country_wide[[col]] / first_value * 100
}

trajectory_summary <- data.frame(
  Country = series_cols,
  Start_value = round(sapply(series_cols, function(col) country_wide[[col]][1]), 1),
  End_value = round(sapply(series_cols, function(col) country_wide[[col]][nrow(country_wide)]), 1),
  Recovery_ratio = round(
    sapply(
      series_cols,
      function(col) country_wide[[col]][nrow(country_wide)] / country_wide[[col]][1]
    ),
    3
  ),
  Indexed_end = round(sapply(series_cols, function(col) country_indexed[[col]][nrow(country_indexed)]), 1)
)

knitr::kable(
  trajectory_summary |>
    dplyr::arrange(dplyr::desc(Recovery_ratio)),
  align = c("l", "r", "r", "r", "r")
)
```

The trajectory summary shows why normalization is necessary. The countries are measured on the same scale, but their growth paths differ enough that raw values alone would overstate the largest markets.

```{r}
country_scaled <- country_indexed
country_scaled[series_cols] <- scale(country_scaled[series_cols])

series_matrix <- t(as.matrix(country_scaled[series_cols]))
rownames(series_matrix) <- series_cols
```

The standardized matrix is now ready for a distance-based clustering model where each row is one country trajectory and each column is one month.

## 6. Diagnose The Number Of Clusters

The first diagnostic is a summary table of candidate `k` values from 2 to 6.

```{r}
cluster_distance <- dist(series_matrix, method = "euclidean")
cluster_tree <- hclust(cluster_distance, method = "ward.D2")

within_cluster_distance <- function(labels, dmat) {
  d <- as.matrix(dmat)
  total <- 0
  for (grp in unique(labels)) {
    idx <- which(labels == grp)
    if (length(idx) > 1) {
      total <- total + sum(d[idx, idx][upper.tri(d[idx, idx])])
    }
  }
  total
}

candidate_k <- 2:6
k_diagnostics <- do.call(
  rbind,
  lapply(candidate_k, function(k) {
    labels <- cutree(cluster_tree, k = k)
    sil <- silhouette(labels, cluster_distance)
    data.frame(
      k = k,
      mean_silhouette = round(mean(sil[, "sil_width"], na.rm = TRUE), 3),
      within_cluster_distance = round(within_cluster_distance(labels, cluster_distance), 1),
      min_cluster_size = min(table(labels))
    )
  })
)

knitr::kable(k_diagnostics, align = c("r", "r", "r", "r"))
```

The diagnostics already point to a compact solution with a small number of clusters. In this dataset, `k = 3` gives the best balance between separation and interpretability.

### 6.1 Diagnostic Plot

The second diagnostic is the silhouette trend across candidate `k` values.

```{r}
#| label: fig-k-diagnostics
ggplot(k_diagnostics, aes(x = k, y = mean_silhouette)) +
  geom_line(color = "#0f3d3e", linewidth = 1.1) +
  geom_point(color = "#d86f45", size = 2.8) +
  scale_x_continuous(breaks = candidate_k) +
  labs(
    title = "Mean Silhouette Score by Candidate k",
    x = "Number of clusters (k)",
    y = "Mean silhouette score"
  ) +
  theme_minimal(base_size = 12)
```

The plot confirms that `k = 3` is the strongest practical choice. It separates the countries cleanly enough to be useful while still leaving room for a meaningful market-state interpretation.

## 7. Fit The Final Hierarchical Model

With the diagnostics in hand, the final clustering model can be fitted using three clusters.

```{r}
selected_k <- 3
final_labels <- cutree(cluster_tree, k = selected_k)

cluster_assignments <- data.frame(
  country = series_cols,
  cluster = paste0("Cluster ", final_labels),
  stringsAsFactors = FALSE
) |>
  dplyr::arrange(cluster, country)

cluster_summary <- cluster_assignments |>
  dplyr::group_by(cluster) |>
  dplyr::summarise(
    cluster_size = dplyr::n(),
    countries = paste(country, collapse = ", "),
    .groups = "drop"
  )

knitr::kable(cluster_summary, align = c("l", "r", "l"))
```

The final solution produces three interpretable blocks: one short-haul Asia-Pacific cluster, one long-haul Europe cluster, and one North Asia cluster centered on China.

## 8. Show Cluster Membership

The next table lists every country and its final cluster assignment.

```{r}
membership_table <- cluster_assignments |>
  dplyr::arrange(cluster, country)

knitr::kable(membership_table, align = c("l", "l"))
```

This membership view is the direct model output that the Shiny app should expose. It lets the user see the grouping instantly and makes the clustering result auditable.

## 9. Show Representative Cluster Profiles

To make the clusters easier to interpret, the next plot shows the average indexed trajectory of each cluster over time, with China highlighted on top of its cluster profile.

```{r}
cluster_profile_long <- do.call(
  rbind,
  lapply(sort(unique(cluster_assignments$cluster)), function(cl) {
    members <- cluster_assignments$country[cluster_assignments$cluster == cl]
    data.frame(
      date = country_indexed$date,
      cluster = cl,
      indexed_value = rowMeans(country_indexed[, members, drop = FALSE], na.rm = TRUE)
    )
  })
)

china_cluster <- cluster_assignments$cluster[cluster_assignments$country == china_series][1]
china_line <- data.frame(
  date = country_indexed$date,
  indexed_value = country_indexed[[china_series]],
  series = china_series
)

ggplot(cluster_profile_long, aes(x = date, y = indexed_value, color = cluster)) +
  geom_line(linewidth = 1.05) +
  geom_line(
    data = china_line,
    aes(x = date, y = indexed_value),
    inherit.aes = FALSE,
    color = "black",
    linewidth = 1.2,
    linetype = "dashed"
  ) +
  scale_color_manual(values = c("Cluster 1" = "#0f6b5b", "Cluster 2" = "#ef9f23", "Cluster 3" = "#4c6faf")) +
  labs(
    title = "Representative Cluster Trajectories with China Highlighted",
    x = "Month",
    y = "Indexed arrivals (first month = 100)",
    color = "Cluster"
  ) +
  theme_minimal(base_size = 12)
```

This plot is the clearest visual summary of the clustering output. It shows the trajectory shape of each cluster and makes it easy to see where China sits relative to its peers.

## 10. Interpret China's Placement

China belongs to the North Asia cluster, together with Japan and Hong Kong SAR (China). The next table shows that cluster and orders the members by their distance to China in the standardized trajectory space.

```{r}
dist_to_china <- as.matrix(cluster_distance)[china_series, ]
china_context <- cluster_assignments |>
  dplyr::filter(cluster == china_cluster) |>
  dplyr::mutate(
    distance_to_china = round(dist_to_china[country], 3)
  ) |>
  dplyr::arrange(distance_to_china)

knitr::kable(china_context, align = c("l", "l", "r"))
```

China is grouped with Japan and Hong Kong SAR (China), which suggests a shared North Asia recovery trajectory rather than a broad global pattern. This is the key interpretive point: China is important, but its recovery shape is more similar to nearby North Asia markets than to the broader Southeast Asia or Europe clusters.

## 11. UI Design Implications

The clustering prototype should expose controls that let the user change the country set, the normalization choice, and the number of clusters without breaking the workflow.

```{r}
ui_map <- data.frame(
  Control = c(
    "Country series selector",
    "Year window",
    "Normalization mode",
    "Cluster count k",
    "Run button"
  ),
  Purpose = c(
    "Choose the series that will be treated as trajectories",
    "Keep the analysis window aligned with the monthly data",
    "Switch between raw, indexed, and z-scored trajectories",
    "Test whether the grouping should be more or less granular",
    "Trigger clustering only when the user is ready"
  ),
  Output = c(
    "Membership table",
    "Comparable time-series matrix",
    "Diagnostic table and profile plot",
    "Cluster summary and China placement",
    "Updated cluster assignments and representative profiles"
  )
)

knitr::kable(ui_map, align = c("l", "l", "l"))
```

This UI mapping is the bridge from the prototype page to the final Shiny module. It tells us exactly which controls need to be exposed and which outputs need to be kept visible for interpretation.
